---
title: "Homework 6"
author: "PSTAT 131/231"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## Tree-Based Models

```{r}
library(tidyverse)
library(tidymodels)
library(rpart.plot)
library(vip)
library(janitor)
library(randomForest)
library(xgboost)
library(ranger)
tidymodels_prefer()
```


### Exercise 1

Read in the data and set things up as in Homework 5:

- Use `clean_names()`
- Filter out the rarer Pokémon types
- Convert `type_1` and `legendary` to factors

Do an initial split of the data; you can choose the percentage for splitting. Stratify on the outcome variable.

Fold the training set using *v*-fold cross-validation, with `v = 5`. Stratify on the outcome variable.

Set up a recipe to predict `type_1` with `legendary`, `generation`, `sp_atk`, `attack`, `speed`, `defense`, `hp`, and `sp_def`:

- Dummy-code `legendary` and `generation`;
- Center and scale all predictors.

```{r}
pokemon <- read.csv("Pokemon.csv")
pokemon_clean <- clean_names(pokemon)

pokemon_filter <- pokemon_clean %>%
  filter(type_1 == c("Bug", "Fire", "Grass", "Normal", "Water", "Psychic")) %>%
  mutate_at(vars(type_1, legendary), factor)

set.seed(2022)

pokemon_split <- initial_split(pokemon_filter, prop = 0.70,
                               strata = type_1)
pokemon_train <- training(pokemon_split)
pokemon_test <- testing(pokemon_split)

dim(pokemon_train)
dim(pokemon_test)

pokemon_folds <- vfold_cv(pokemon_train, v = 5, strata = type_1)
pokemon_folds

pokemon_recipe <- recipe(type_1 ~ legendary + generation + sp_atk + attack
                         + speed + defense + hp + sp_def, data = pokemon_train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_center(all_numeric_predictors()) %>%
  step_scale(all_numeric_predictors())

pokemon_recipe

```


### Exercise 2

Create a correlation matrix of the training set, using the `corrplot` package. *Note: You can choose how to handle the continuous variables for this plot; justify your decision(s).*

What relationships, if any, do you notice? Do these relationships make sense to you?

```{r}
library(corrplot)

cor_pokemon <- pokemon_train %>%
  select(c(generation, sp_atk, attack, speed, defense, hp, sp_def)) %>%
  cor()

corrplot(cor_pokemon, method = "number")

```


### Exercise 3

First, set up a decision tree model and workflow. Tune the `cost_complexity` hyperparameter. Use the same levels we used in Lab 7 -- that is, `range = c(-3, -1)`. Specify that the metric we want to optimize is `roc_auc`. 

Print an `autoplot()` of the results. What do you observe? Does a single decision tree perform better with a smaller or larger complexity penalty?

```{r}
tree_spec <- decision_tree() %>%
  set_engine("rpart")

class_tree_spec <- tree_spec %>%
  set_mode("classification")

class_tree_wf <- workflow() %>%
  add_model(class_tree_spec %>% set_args(cost_complexity = tune())) %>%
  add_recipe(pokemon_recipe)

param_grid <- grid_regular(cost_complexity(range = c(-3, -1)), levels = 10)

tune_res <- tune_grid(
  class_tree_wf,
  resamples = pokemon_folds,
  grid = param_grid,
  metrics = metric_set(roc_auc)
)

autoplot(tune_res)

```

Solution: A single decision tree performs better with a smaller complexity penalty.

### Exercise 4

What is the `roc_auc` of your best-performing pruned decision tree on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

```{r}
collect_metrics(tune_res) %>%
  arrange(-mean)
```

Solution: The roc_auc of best-performing pruned decision tree on the folds is 0.651667.

### Exercise 5

Using `rpart.plot`, fit and visualize your best-performing pruned decision tree with the *training* set.

```{r}
best_complexity <- select_best(tune_res, metric = "roc_auc")

class_tree_final <- finalize_workflow(class_tree_wf, best_complexity)

class_tree_final_fit <- fit(class_tree_final, data = pokemon_train)

class_tree_final_fit %>%
  extract_fit_engine() %>%
  rpart.plot()
```


### Exercise 5

Now set up a random forest model and workflow. Use the `ranger` engine and set `importance = "impurity"`. Tune `mtry`, `trees`, and `min_n`. Using the documentation for `rand_forest()`, explain in your own words what each of these hyperparameters represent.

Create a regular grid with 8 levels each. You can choose plausible ranges for each hyperparameter. Note that `mtry` should not be smaller than 1 or larger than 8. **Explain why not. What type of model would `mtry = 8` represent?**

```{r}
rf_spec <- rand_forest() %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("classification")

rf_wf <- workflow() %>% 
  add_model(rf_spec %>% set_args(mtry = tune(), trees = tune(), min_n = tune())) %>%
  add_recipe(pokemon_recipe)

rf_grid <- grid_regular(mtry(range = c(1, 8)), trees(range = c(1, 8)), 
                        min_n(range = c(1, 8)), levels = 8)


```


Solution: mtry: the number of predictors that will be randomly sampled.

trees: the number of trees contained in the ensemble.

min_n: the minimum number of data points in a node that are required for the node to be split further.

Because there are only 8 predictors thus 1 <= mtry <= 8. mtry = 8 means we use all the predictors to be randomly sampled at each split when creating the tree models.

### Exercise 6

Specify `roc_auc` as a metric. Tune the model and print an `autoplot()` of the results. What do you observe? What values of the hyperparameters seem to yield the best performance?

```{r}
rf_tune_res <- tune_grid(
  rf_wf,
  resamples = pokemon_folds,
  grid = rf_grid,
  metrics = metric_set(roc_auc)
)

autoplot(rf_tune_res)
```

Solution: The relationship is complex. trees = 8, mtys = 5, min_n = 7 seems to yield the best performance.

### Exercise 7

What is the `roc_auc` of your best-performing random forest model on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

```{r}
collect_metrics(rf_tune_res) %>%
  arrange(-mean)
```

Solution: The roc_auc of your best-performing random forest model on the folds is 0.7114815.

### Exercise 8

Create a variable importance plot, using `vip()`, with your best-performing random forest model fit on the *training* set.

```{r}
best_complexity <- select_best(rf_tune_res, metric = "roc_auc")

rf_final <- finalize_workflow(rf_wf, best_complexity)

rf_final_fit <- fit(rf_final, data = pokemon_train)

rf_final_fit %>%
  pull_workflow_fit() %>%
  vip()
```


Which variables were most useful? Which were least useful? Are these results what you expected, or not?

Solution: sp_atk is most useful and legendary is least useful. Yes, the sp_atk is the most important character of the type of pokemons.

### Exercise 9

Finally, set up a boosted tree model and workflow. Use the `xgboost` engine. Tune `trees`. Create a regular grid with 10 levels; let `trees` range from 10 to 2000. Specify `roc_auc` and again print an `autoplot()` of the results. 

What do you observe?

What is the `roc_auc` of your best-performing boosted tree model on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

```{r}
boost_spec <- boost_tree() %>%
  set_engine("xgboost") %>%
  set_mode("classification")

boost_wf <- workflow() %>% 
  add_model(boost_spec %>% set_args(trees = tune())) %>%
  add_recipe(pokemon_recipe)

boost_grid <- grid_regular(trees(range = c(10, 2000)), levels = 10)

boost_tune_res <- tune_grid(
  boost_wf,
  resamples = pokemon_folds,
  grid = boost_grid,
  metrics = metric_set(roc_auc)
)

autoplot(boost_tune_res)

collect_metrics(boost_tune_res) %>%
  arrange(desc(mean), by_group = TRUE)
```

Solution: We observe that with the increase of the trees, the roc_auc increases quickly and slow down and last finally. The roc_auc of the best-performing boosted tree model on the folds is 0.6492593.

### Exercise 10

Display a table of the three ROC AUC values for your best-performing pruned tree, random forest, and boosted tree models. Which performed best on the folds? Select the best of the three and use `select_best()`, `finalize_workflow()`, and `fit()` to fit it to the *testing* set. 

Print the AUC value of your best-performing model on the testing set. Print the ROC curves. Finally, create and visualize a confusion matrix heat map.

Which classes was your model most accurate at predicting? Which was it worst at?

```{r}
roc_auc_pre <- c(0.6516667, 0.7114815, 0.6492593)
models <- c("best-performing pruned tree model", "random forest model", "boosted tree model")
results <- tibble(roc_auc_pre = roc_auc_pre, models = models)

results %>%
  arrange(-roc_auc_pre)

best_complexity <- select_best(rf_tune_res, metric = "roc_auc")

rf_final <- finalize_workflow(rf_wf, best_complexity)

rf_final_fit <- fit(rf_final, data = pokemon_train)

augment(rf_final_fit, new_data = pokemon_test) %>%
  roc_auc(type_1, .pred_Bug:.pred_Water) 

augment(rf_final_fit, new_data = pokemon_test) %>%
  roc_curve(type_1, .pred_Bug:.pred_Water) %>%
  autoplot()

augment(rf_final_fit , new_data = pokemon_test) %>%
  conf_mat(type_1, .pred_class) %>%
  autoplot(type = "heatmap")

```

Solution: The random forest model performed best on the folds. The model is best at predicting fire and is worst at predicting grass.

